---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
```

# Tariff Data

BMW 1er (2020)

```{r}
tariff_df <- tibble(
    Name = c('1M', '1H', '2H', '4H', '6H', '9H', '1D', '2D', '3D', '4D', '5D', '6D', '7D', '8D', '9D', '10D', '11D', '12D', '13D', '14D', '15D', '16D', '17D', '18D', '19D', '20D', '21D', '22D', '23D', '24D', '25D', '26D', '27D', '28D', '29D', '30D'),
    GP = c(rep(0.99, 6), rep(0, 30)),
    TP = c(0.4, 16.99, 22.99, 35.49, 47.99, 62.99, 69.99, 134.99, 189.99, 239.99, 279.99, 314.99, 349.99, 384.99, 419.99, 454.99, 489.99, 524.99, 599.99, 594.99, 629.99, 659.99, 689.99, 714.99, 744.99, 469.99, 799.99, 829.99, 854.99, 884.99, 909.99, 939.99, 969.99, 994.99, 1024.99, 1049.99),
    ExtraMinsPrice = c(0.4, rep(0.36, 35)),
    DailyCap = c(rep(99.99, 6), rep(82.99, 3), rep(61.99, 36-6-3)),
    PrepaidKm = c(200, rep(0, 35)),
    ExtraKmPrice = c(0.39, rep(0.23, 35))
) |> 
    mutate(
        pn = str_extract(Name, '[:digit:]+') |> as.numeric(),
        PrepaidMins = case_match(
            str_extract(Name, '[:alpha:]'),
            'M' ~ pn,
            'H' ~ pn * 60,
            'D' ~ pn * 60 * 24
        ),
        .after = 'TP'
    ) |> 
    select(-pn)
tariff_df
```

-   Prices: EUR

-   Duration: minutes

-   Distance: kilometres

Possible inaccuracies:

-   For 1M: 0.23 per extra km after 99.99 cap (?)

-   For 1M: 0.36 per extra km on subsequent days (?)

# `TravelPrice` Function

$$
\mathrm{TravelPrice}(t, s) = \mathrm{GP} + \mathrm{TP} + \min(\mathrm{ExtraDuration}(t)\cdot \mathrm{ExtraMinPrice}, \mathrm{DailyCap}) + \mathrm{ExtraDist}(s)\cdot \mathrm{ExtraKmPrice}
$$

```{r}
extra <- function(x, prepaid) {
    pmax(0, x - prepaid)
}
stopifnot(extra(c(0, 23), 10) == c(0, 13))
```

-   $\mathbf{t}$, $\mathbf{s}$: vectors

-   $\boldsymbol{\theta}$: tariff parameter vector

```{r}
travel_price <- function(t, s, θ, relative = FALSE) {
    p <- θ$GP + θ$TP + 
        pmin(extra(t, θ$PrepaidMins) * θ$ExtraMinsPrice, θ$DailyCap) +
             extra(s, θ$PrepaidKm)   * θ$ExtraKmPrice
    if (relative) {
        p <- p / t / s
    }
    p
}
```

Travel 90, 120 minutes and 70, 80 kilometres

```{r}
travel_price(c(90, 120), c(70, 80), tariff_df[2,])
travel_price(c(90, 120), c(70, 80), tariff_df[2,], relative = TRUE)
```

# Distance x Duration

Distance x Duration grid

```{r}
make_grid <- function(
        min_duration = 1, max_duration,
        min_distance = 1, max_distance, 
        # km/m
        min_speed = 1, max_speed
    ) {
    crossing(
        duration = min_duration:max_duration,
        distance = min_distance:max_distance
    ) |> 
    filter(between(distance / duration, min_speed, max_speed))
}

max_duration <- 60 * 12
max_distance <- 1000
min_speed <- 15 / 60
max_speed <- 150 / 60

td_grid <- make_grid(
    max_duration = max_duration,
    max_distance = max_distance,
    min_speed = min_speed, max_speed = max_speed
)
    
td_grid |> nrow()
```

Scalar fields: absolute price

```{r}
1:6 |> 
    map(\(k) {
        params <- tariff_df[k,]
        td_grid |> 
            filter(duration > params$PrepaidMins) |>
            mutate(RTP = travel_price(duration, distance, params)) |> 
            ggplot(aes(x = duration, y = distance, z = RTP)) +
            geom_contour_filled(bins = 10) +
            labs(title = params$Name)
    })
```

relative price (/km/min)

```{r}
1:6 |> 
    map(\(k) {
        params <- tariff_df[k,]
        td_grid |> 
            filter(duration > params$PrepaidMins) |>
            mutate(RTP = travel_price(duration, distance, params, relative = TRUE)) |> 
            ggplot(aes(x = duration, y = distance, z = RTP)) +
            geom_contour_filled(bins = 10) +
            labs(title = params$Name)
    })
```

# Optimisation Across Tariffs

5 Hours, 200 kilometers

```{r}
1:6 |> 
    map_dfr(\(k) {
        params <- tariff_df[k,]
        tibble(
            Name = params$Name,
            Price = travel_price(5 * 60, 200, params)
            # `Price/Km/Min` = travel_price(5 * 60, 200, params, relative = TRUE)
        )
    })
```

```{r}
min_tariff_df <- 
    1:6 |> 
    map(\(k) {
        params <- tariff_df[k,]
        td_grid |> 
            # filter(duration > params$PrepaidMins) |>
            mutate(
                RTP = travel_price(duration, distance, params, relative = TRUE),
                TariffName = params$Name
            ) 
    }) |> 
    list_rbind() |> 
    pivot_wider(names_from = 'TariffName', values_from = 'RTP') |> 
    mutate(
        min_idx = (pick(-(1:2)) |> as.matrix() |> apply(1, which.min)),
        min_tariff_name = names(pick(-(1:2)))[min_idx]
    )
plot_min_tariff <- function(df, font_size = 9) {
    df |> 
        mutate(
            duration = duration / 60,
            min_tariff_name = relevel(as.factor(min_tariff_name), '1M')
        ) |> 
        ggplot(aes(x = duration, y = distance)) +
        geom_hex(aes(fill = min_tariff_name), stat = 'identity') +
        scale_fill_discrete(name = 'Tariff Name') +
        labs(
            x = 'Duration (hr)', y = 'Distance (km)', 
            title = 'Most Favourable ShareNow Tariffs',
            subtitle = 'BMW 1er (2020)',
            caption = glue::glue('Min. speed = {min_speed * 60} km/h; max. speed = {max_speed * 60} km/h')
        ) +
        scale_x_continuous(breaks = 0:(max_duration/60)) +
        scale_y_continuous(breaks = seq(0, max_distance, by = 100)) +
        theme_bw(base_size = font_size)
}
min_tariff_df |> 
    # group_by(duration) |>
    # slice_sample(n = 100) |> 
    # ungroup() |> 
    plot_min_tariff()
min_tariff_df |> 
    filter(distance <= 200) |> 
    plot_min_tariff()
```

```{r}
pdf('./romas_card-all-a7.pdf', width = 4.1, height = 2.9)
min_tariff_df |> plot_min_tariff(font_size = 7)
dev.off()
```

```{r}
pdf('./romas_card-all-a6.pdf', width = 5.8, height = 4.1)
min_tariff_df |> plot_min_tariff(font_size = 9)
dev.off()
```

```{r}
pdf('./romas_card-closeup-a6.pdf', width = 5.8, height = 4.1)
min_tariff_df |> filter(distance <= 200) |> plot_min_tariff(font_size = 9)
dev.off()
```

```{r}
pdf('./romas_card-closeup-a7.pdf', width = 4.1, height = 2.9)
min_tariff_df |> filter(distance <= 200) |> plot_min_tariff(font_size = 7)
dev.off()
```

# Fixed Duration

5 Hours

```{r}
target_duration <- 5 * 60
dist_df <- 
    1:6 |> 
    map_dfr(\(k) {
        params <- tariff_df[k,]
        tibble(
            Name = params$Name,
            Duration = target_duration,
            Distance = (10 * 5):(100 * 5)
        ) |> 
            mutate(
                Price = travel_price(Duration, Distance, params),
                `Price/Km/Min` = travel_price(Duration, Distance, params, relative = TRUE)
            )
    })
dist_df |> 
    ggplot() +
    geom_line(aes(x = Distance, y = `Price`, color = Name)) +
    labs(title = glue::glue('Target duration: {target_duration / 60} hours'))
dist_df |> 
    ggplot() +
    geom_line(aes(x = Distance, y = `Price/Km/Min`, color = Name)) +
    labs(title = glue::glue('Target duration: {target_duration / 60} hours'))
```

# Fixed Distance

200 Kilometers

```{r}
target_distance <- 200
dist_df <- 
    1:6 |> 
    map_dfr(\(k) {
        params <- tariff_df[k,]
        tibble(
            Name = params$Name,
            Distance = target_distance,
            Duration = (2*60):(6*60)
        ) |> 
            mutate(
                Price = travel_price(Duration, Distance, params),
                `Price/Km/Min` = travel_price(Duration, Distance, params, relative = TRUE),
                DurationH = Duration / 60
            )
    })
dist_df |> 
    ggplot() +
    geom_line(aes(x = DurationH, y = `Price`, color = Name)) +
    labs(title = glue::glue('Target distance: {target_distance} km'))
dist_df |> 
    ggplot() +
    geom_line(aes(x = DurationH, y = `Price/Km/Min`, color = Name)) +
    labs(title = glue::glue('Target distance: {target_distance} km'))
```
